{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2017-01-05 proposal/410.tex
timeline
write about education implementation

>>>>

\section{Motivation}

In my MS thesis \cite{landslide}, I solicited students at CMU's Operating Systems Design and Implementation class
(henceforth, ``15-410'')
to volunteer at the end of the P3 project to annotate their kernels and try debugging them with Landslide.
However, the annotation burden undermined Landslide's purpose:
the only students willing to spend free time on manual instrumentation were biased to be those who were already doing well in the class,
and hence least likely to benefit from Landslide's debugging potential.
(Actually, even the best 15-410 students still have concurrency bugs,
but in principle, an educational tool must reach the more struggling students,
the so-called ``middle'' and/or ``bottom'' of the class.)
%
Requiring annotations hurts Landslide's case as a grading tool, as well:
TAs need to understand the kernel to begin with in order to annotate correctly,
and while achieving such understanding they may as well grade it by hand, as before.

Since then, I've extended Landslide to support testing P2 thread libraries (Section~\ref{sec:pebbles}) as well.
Because P2 mandates specific function names for the project's internal APIs
-- most importantly, for the concurrency primitives --
Landslide can automatically annotate arbitrary student implementations with no manual effort required of the user (whether student or TA).
The addition of Quicksand and Iterative Deepening (Chapter~\ref{chap:quicksand}) partly fulfills this purpose,
freeing student attention from the issue of which state spaces to test.
This chapter will detail my further techniques and evaluation which are specific to educational use.

\section{Implementation Details}

Every stateless model checker must make some assumptions about the tested programs' concurrency model \cite{chess,inspect,dbug-ssv}.
However, arbitrary programs may break conventional disciplines of concurrent programs, while still being bug-free.
For example, thread communication via ad-hoc {\tt yield} loops may appear as an infinite loop, livelock, or deadlock.
This is especially true of student code,
written by people who are just learning concurrent programming discipline for the first time,
and/or written under the time pressure of a project deadline.
To be an effective tool for struggling students,
a model checker must somehow coax adversarial programs to fit its concurrency model, to effectively test for real bugs,
rather than rejecting them outright on some stylistic or disciplinary grounds.

Fully-automatic instrumentation of student P2s has been no walk in the park.
I have equipped Landslide with several powerful algorithms and heuristics for handling the most common anti-patterns in student submissions.

% TODO
\begin{itemize}
	\item 
	\item 
	\item 
\end{itemize}

However, some of the less common offenses are both more difficult to handle algorithmically, and also more worrying from a pedagogical point of view.
%After some collaboration with David Eckhardt (15-410 professor and member of this thesis committee),
For the following cases, we configured Landslide to abort, and warn the student that they must find a better solution before it could test their code.

\begin{itemize}
	\item Busy-wait loops containing neither {\tt yield} nor {\tt xchg} (nor any other atomic instruction), such as {\tt while (!other\_thread\_ready) continue;}.
		This blurs the line between anti-pattern and concurrency bug:
		because it does not yield the CPU, a uni-processor machine must wait for the next timer tick (several milliseconds!) before making any progress;
		also, because it does not use atomic instructions, an optimizing compiler may reorder or even delete the loop's accesses.
		%
		Landslide also cannot easily identify it as similar to message-passing,
		appearing indistinguishable
		% not really? there would be dpor conflix, at least
		from a thread-local infinite computation,
		which is of course impossible to judge for halting \cite{entscheidungsproblem}.

		In such a case, Landslide will issue a bug report with the special message:
		{\em I have run a loop in [function name] an alarming number of times.
		This version of Landslide cannot distinguish between this loop being infinite versus merely undesirable.
		Please refer to the ``Synchronization (2)'' lecture.}
	\item Recursive mutex use (i.e., locking the same mutex twice in the same thread, then subsequently unlocking it twice).
		While it would not be difficult for Landslide's lock-sets to support recursive locking (using a nesting counter instead of a boolean flag),
		that would assume the corresponding mutex implementation provides the same semantics,
		which is risky business with student code. % ;)
		Furthermore, recursive locking is not an obvious solution to any of P2's challenges;
		far more often, it arises when a student's mutex tries to {\tt malloc} some internal state,
		which itself requires a mutex for safe allocation, which can lead to stack overflow and a crash.

		In such a case, Landslide will issue a bug report with the special message:
		{\em This version of Landslide cannot debug recursive implementations of mutex\_lock.
		Please examine this stack trace and determine for yourself whether it indicates a bug.}
\end{itemize}

\section{Landslide as a Debugging Tool}
\label{sec:studence}

TODO

\section{Landslide as a Grading Tool}
\label{sec:grading}

TODO
